### 2019-02-25 R Files/data_cleaning.Rmd

---
title: "Data_Cleaning"
author: "Aidan Mullan"
date: "2/5/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
## Read in the raw data files
chest <- read_excel("Chest Only.xlsx")
CTA <- read_excel("CTA.xlsx")

## Remove unnecessary variables (descriptions, repeated columns, etc)
chest_short <- chest[1:2413,c(1:6,8:9,11:16)]
CTA_short <- CTA[,c(1:6,8:9,11:16)]

## Combine chest and CTA datasets
names(CTA_short) <- names(chest_short)
PE_raw <- rbind(chest_short, CTA_short)
```

```{r}
## Collapse factors into main categories for 'Procedure' and 'CT Type'
pro_index <- which(PE_raw$Procedure %in% c("Triple rule out", NA))
PE2 <- PE_raw[-pro_index,]
PE2$Procedure <- ifelse(PE2$Procedure == "Chest only", "Chest Only",
                        ifelse(PE2$Procedure %in% c("Chest and Abdomen", "Neck and Chest",
                                                    "Chest, Abdomen and Pelvis"), "CAP", PE2$Procedure))

PE2$`CT Type` <- ifelse(PE2$`CT Type` %in% c("DS_Force", "DS_Drive"), "DS_Scanner", PE2$`CT Type`)

## Give consistency to 'Location', 'Pulmonary Embolism', 'Type', and 'Strain' factor names
PE2$Location <- ifelse(PE2$Location == "Out", "OUT", 
                       ifelse(PE2$Location == "In", "IN", PE2$Location))

PE2$`Pulmonary Embolism` <- ifelse(PE2$`Pulmonary Embolism` == "No", "NO", 
                       ifelse(PE2$`Pulmonary Embolism` == "Yes", "YES", PE2$`Pulmonary Embolism`))

PE2$Type <- ifelse(PE2$Type == "peripheral", "Peripheral", PE2$Type)
PE2$Type <- ifelse(is.na(PE2$Type), "None", PE2$Type)

PE2$Strain <- ifelse(PE2$Strain %in% c('Yes', 'yes', 'YEs'), "YES",
                     ifelse(PE2$Strain == "No", "NO", PE2$Strain))
PE2$Strain <- ifelse(is.na(PE2$Strain), "None", PE2$Strain)

PE2$BMI[c(112,264,416,453,1876,2188,2217,2371,2650)] <- NA
PE2$BMI <- ifelse(PE2$BMI == "None", NA, PE2$BMI)
```

```{r}
## Convert variables to factors with specified levels
PE2$Gender <- as.factor(PE2$Gender); PE2$Procedure <- as.factor(PE2$Procedure)
PE2$`CT Type` <- as.factor(PE2$`CT Type`); PE2$Location <- as.factor(PE2$Location)
PE2$`Pulmonary Embolism` <- as.factor(PE2$`Pulmonary Embolism`)
PE2$Type <- as.factor(PE2$Type); PE2$Strain <- as.factor(PE2$Strain)
PE2$BMI <- as.numeric(PE2$BMI); PE2$`Contrast Given` <- as.numeric(PE2$`Contrast Given`)

## Construct age/BMI grouping variables
PE2$Age_range <- cut(PE2$Age, breaks = c(0,18,35,110), labels = c("<18", "18-35", ">35"))
PE2$BMI_range <- cut(PE2$BMI, breaks = c(0,25,40,100), labels = c("<25", "25-40", ">40"))
```

```{r}
write.csv(PE2, "PE_clean.csv")
```