### 2019-02-27 R Files/Update_Analysis.Rmd
Began analysis on the updated dataset

---
title: "Analysis2"
author: "Aidan Mullan"
date: "2/26/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
PE <- read_excel("Data/Updates/Book2.xlsx")
PE$BMI <- as.numeric(PE$BMI)
PE <- na.omit(PE)
summary(PE)
```

```{r}

## Looking at Contrast: Histogram and Boxplot
par(mfrow = c(1,2))
hist(PE$`Contrast Given`, breaks = 20, main = "Figure 2a: Histogram of Contrast Given",
     xlab = "Contrast Given", ylab = "No. Patients", cex.main = 0.75)
boxplot(PE$`Contrast Given`, main = "Figure 2b: Boxplot of Contrast Given", cex.main = 0.75, 
        ylab = "Contrast Given", yaxp = c(0,150,10))
par(mfrow = c(1,1))

## Looking at Effective Dose: Histogram and Boxplot
par(mfrow = c(1,2))
hist(PE$`Eff Dose`, breaks = 20, main = "Figure 1a: Histogram of Effective Dose",
     xlab = "Effective Dose", ylab = "No. Patients", cex.main = 0.75)
axis(side = 1, at = seq(0,100,10))
boxplot(PE$`Eff Dose`, main = "Figure 1b: Boxplot of Effective Dose", cex.main = 0.75, 
        ylab = "Effective Dose")
axis(side = 2, at = seq(0,100,10))
par(mfrow = c(1,1))
```

```{r}
## Plotting Contrast by Age (10 Year Intervals)
age_cuts <- c("<10", "10-20", "20-30", "30-40", "40-50","50-60", "60-70",
              "70-80", "80-90", "90-100", ">100")
PE$Age10 <- cut(PE$Age, breaks = seq(0,110,10), labels = age_cuts)
summary(PE$Age10)

contrast_by_age <- tapply(PE$`Contrast Given`, PE$Age10, "mean")
plot(contrast_by_age, type = "l", ylab = "Contrast Given", xlab = "Age Group", xaxt = "n",
     main = "Avergae Contrast Administered by Age")
points(contrast_by_age)
axis(1, at = 1:11, labels = age_cuts, cex.axis = .75)

## Plotting Radiation by Age (10 Year Intervals)
rad_by_age <- tapply(PE$`Eff Dose`, PE$Age10, "mean")
plot(rad_by_age, type = "l", ylab = "Effective Dose", xlab = "Age Group", xaxt = "n",
     main = "Mean Effective Dose by Age")
points(rad_by_age)
axis(1, at = 1:11, labels = age_cuts, cex.axis = .75)

## Plotting Contrast by BMI (5 Unit Intervals)
BMI_cuts <- c("10-15", "15-20", "20-25", "25-30","30-35", "35-40",
              "40-45", "45-50", "50-55", "55-60", ">60")
PE$BMI10 <- cut(PE$BMI, breaks = c(seq(10,60,5),92), labels = BMI_cuts)
summary(PE$BMI10)

contrast_by_bmi <- tapply(PE$`Contrast Given`, PE$BMI10, "mean")
plot(contrast_by_bmi, type = "l", ylab = "Contrast Given", xlab = "BMI Group", xaxt = "n",
     main = "Avergae Contrast Administered by BMI")
points(contrast_by_bmi)
axis(1, at = 1:11, labels = BMI_cuts, cex.axis = .75)

## Plotting Radiation by BMI (5 Unit Intervals)
rad_by_bmi <- tapply(PE$`Eff Dose`, PE$BMI10, "mean")
plot(rad_by_bmi, type = "l", ylab = "Effective Dose", xlab = "BMI Group", xaxt = "n",
     main = "Mean Effective Dose by BMI")
points(rad_by_bmi)
axis(1, at = 1:11, labels = BMI_cuts, cex.axis = .75)
```