### 2019-02-27 R Files/Updates_Writeup.tex

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Update\_Writeup},
            pdfauthor={Aidan Mullan},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Update\_Writeup}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Aidan Mullan}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{2/26/2019}

\usepackage{float}

\begin{document}
\maketitle

\begin{verbatim}
## Warning in read_fun(path = path, sheet_i = sheet, limits = limits, shim =
## shim, : Expecting numeric in I1195 / R1195C9: got 'l00'
\end{verbatim}

\begin{verbatim}
## Warning: NAs introduced by coercion
\end{verbatim}

\subsection{Effective Dose}\label{effective-dose}

\paragraph{Summary Statistics}\label{summary-statistics}

Using the fixed data for effective dosage of radiation administered to
patients, we retrieve the plots shown in Figure 1. Here we can see the
right-skew in the histogram suggesting several patients recieved much
higher amounts of radiation than average. The mean effective dose given
to a patient was 5.631 millisieverts, with a range of 0.1 to 48.8
millisieverts. Out of all 2382 patients, only 54 recieved a radiation
dose above 20 millisieverts.

\begin{table}[H] \centering 
  \caption{Summary Statistics of Radiation and Contrast} 
\begin{tabular}{p{3cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}}
\\[-1.8ex] \hline 
\hline
 & Min & 1st Q & Median & Mean & 3rd Q & Max \\ 
\hline \\[-1.8ex] 
Effective Dose & 0.10 & 3.06 & 4.30 & 5.631 & 6.36 & 48.83 \\
Contrast       & 5.0 & 75.0 & 95.0 & 88.5 & 98.0 & 150.0 \\
\hline 
\hline \\[-1.8ex]
 \end{tabular}
\end{table}

\includegraphics{Updates_Writeup_files/figure-latex/unnamed-chunk-2-1.pdf}

\begin{verbatim}
##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
##   0.098   3.055   4.298   5.631   6.356  48.832
\end{verbatim}

We are also interested in the relationship between the effecive dose
administered and both the age and BMI of the patient. For the case of
age, patients were grouped into 10-year age groups. The number of
patients that belong to group is given below. Figure 2 shows the average
radiation administered for each of these age groups.

\begin{verbatim}
##   <20 20-30 30-40 40-50 50-60 60-70 70-80 80-90   >90 
##    55   133   222   294   514   562   397   179    26
\end{verbatim}

\includegraphics{Updates_Writeup_files/figure-latex/unnamed-chunk-3-1.pdf}

\subsection{Contrast Given}\label{contrast-given}

\end{document}