### 2019-02-27 R Files/Update_Cleaning.Rmd

---
title: "Update_Cleaning"
author: "Aidan Mullan"
date: "2/27/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
PE_raw <- read_excel("Data/Updates/chest_only.xlsx")
names(PE_raw)

PE_short <- PE_raw[1:2413,c(3,4,6,8,9,11,12,13)]
unique(PE_short$Location)

PE_short$Location <- ifelse(PE_short$Location == "Out", "OUT", 
                       ifelse(PE_short$Location == "In", "IN", PE_short$Location))
PE_short$BMI <- as.numeric(PE_short$BMI)
PE_short$`Contrast Given` <- as.numeric(PE_short$`Contrast Given`)
PE_short$`CT Type` <- ifelse(PE_short$`CT Type` %in% c("DS_Force", "DS_Drive"),
                             "DS_Scanner", PE_short$`CT Type`)
PE_short$Procedure <- ifelse(PE_short$Procedure == "Chest only", "Chest Only", PE_short$Procedure)

PE_chest <- subset(PE_short, subset = Procedure == "Chest Only")
PE <- na.omit(PE_chest)

write.csv(PE, file = "PE_update_clean.csv")
```