### 2019-03-08 R Files/Updated Model/Contrast_Model.Rmd
Re-re-fit the contrast model and began writeup

---
title: "Contrast_Model"
author: "Aidan Mullan"
date: "3/7/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
PE_raw <- read.csv("Data/1st_Set/PE_clean.csv")
PE_short <- PE_raw[,c(4,5,8,9,10,11,12)]
PE_chest <- subset(PE_short, subset = Procedure == "Chest Only")

index <- which(PE_chest$Contrast.Given > 150)
PE <- na.omit(PE_chest[-index,])
```

```{r}
PE$Age_range <- cut(PE$Age, breaks = c(0,18,35,110), labels = c("<18", "18-35", ">35"))
PE$BMI_range <- cut(PE$BMI, breaks = c(0,25,40,100), labels = c("<25", "25-40", ">40"))

contrast1 <- lm(Contrast.Given~CT.Type+BMI_range+Age_range+Gender+Location, data = PE)
summary(contrast1)

contrast2 <- update(contrast1, .~.-Location)
summary(contrast2)
anova(contrast1, contrast2) # p = .026, prefer full model

contrast.aov <- aov(Contrast.Given~CT.Type+BMI_range+Age_range+Gender+Location, data = PE)
TukeyHSD(contrast.aov, which = "CT.Type")
```