### 2019-07-18 R Files/Manuscript_Graphics.Rmd
Image updates

---
title: "Manuscript_Graphics"
author: "Aidan Mullan"
date: "6/22/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(tidyverse)
PE_oldraw <- read.csv("Data/1st_Set/PE_clean.csv")
PE_old <- na.omit(subset(PE_oldraw[,c(3,4,5,7:13,16,17)], subset = Procedure == "Chest Only"))
```

```{r}

## Patient Age and BMI Demographics
counts <- as.data.frame(table(PE_oldraw$Age_range, PE_oldraw$BMI_range))
names(counts) <- c("Age", "BMI", "n")
counts$Age <- factor(counts$Age, levels = c("<18", "18-35", ">35"))
counts$BMI <- factor(counts$BMI, levels = c("<25", "25-40", ">40"))

age_bmi <- ggplot(counts, aes(fill = BMI, x = Age, y = n, label = n)) + 
              geom_bar(position = "dodge", stat = "identity") +
              scale_fill_manual(values = c("lightblue", "dodgerblue", "navyblue")) +
              scale_y_continuous(breaks = seq(0,1500,250)) + 
              ylab("Number of Patients") + 
              ggtitle("Patient Population by Age and BMI Group")

bar_AgeBMI <- ggplot(counts, aes(fill = BMI, x = Age, y = n, label = n)) + 
              geom_bar(position = "dodge", stat = "identity") +
              geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
              scale_fill_manual(values = c("lightblue", "dodgerblue", "navyblue")) +
              scale_y_continuous(breaks = seq(0,1500,250)) + 
              ylab("Number of Patients") + 
              ggtitle("Patient Population by Age and BMI Group")

ggsave(filename = "Images/AgeBMI.jpeg", plot = age_bmi,
       width = 7, height = 5, device = "jpeg")
ggsave(filename = "Images/Bar_AgeBMI.jpeg", plot = bar_AgeBMI,
       width = 7, height = 5, device = "jpeg")
```

```{r}
#####--------------------
## Observed Bar Plots
#####--------------------

PE_old$Positivity <- if_else(PE_old$Pulmonary.Embolism == "YES", 1, 0)

## Positivity X age, bmi

Pos_Age <- PE_old %>% group_by(Age_range) %>% summarise(n = n(), Positivity = mean(Positivity))
Pos_Age$Positivity <- round(Pos_Age$Positivity, 3)
Pos_Age$Age_range <- factor(Pos_Age$Age_range, levels = c("<18", "18-35", ">35"))
bar_PosAge <- ggplot(Pos_Age, aes(fill = Age_range, x = Age_range, y = Positivity, label = Positivity)) +
                     ylim(0,0.25) +
                     geom_bar(position = "dodge", stat = "identity", width = 0.5) +
                     geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
                     scale_fill_manual(values = c("lightblue", "dodgerblue" ,"navyblue")) +
                     ylab("Pulmonary Embolism Positivity Rate") + 
                     xlab("Age Group") +
                     labs(fill = "Age Group") +
                     theme(legend.position = "none") +
                     ggtitle("Pulmonary Embolism Positivity Rate by Age")

Pos_BMI <- PE_old %>% group_by(BMI_range) %>% summarise(n = n(), Positivity = mean(Positivity))
Pos_BMI$Positivity <- round(Pos_BMI$Positivity, 3)
Pos_BMI$Signif <- c("*", "", "*")
Pos_BMI$BMI_range <- factor(Pos_BMI$BMI_range, levels = c("<25", "25-40", ">40"))
bar_PosBMI <- ggplot(Pos_BMI, aes(fill = BMI_range, x = BMI_range, y = Positivity, label = Positivity)) + 
                     ylim(0,0.25) +
                     geom_bar(position = "dodge", stat = "identity", width = 0.5) +
                     geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
                     geom_text(aes(label = Signif), size = 5, position = position_dodge(0.9),
                               vjust = -1.5) +
                     scale_fill_manual(values = c("lightblue", "dodgerblue", "navyblue")) +
                     ylab("Pulmonary Embolism Positivity Rate") + 
                     xlab("BMI Group") +
                     labs(fill = "BMI Group") +
                     theme(legend.position = "none") + 
                     ggtitle("Pulmonary Embolism Positivity Rate by BMI")

ggsave(filename = "Images/Bar_PosAge.jpeg", plot = bar_PosAge,
       width = 7, height = 5, device = "jpeg")
ggsave(filename = "Images/Bar_PosBMI.jpeg", plot = bar_PosBMI,
       width = 7, height = 5, device = "jpeg")

## Positivity X age+gender, bmi+gender

Pos_AgeGender <- PE_old %>% group_by(Age_range, Gender) %>% summarise(n = n(),
                                                                      Positivity = mean(Positivity))
Pos_AgeGender$Positivity <- round(Pos_AgeGender$Positivity, 3)
Pos_AgeGender$Signif <- c("","","","","*","*")
Pos_AgeGender$Age_range <- factor(Pos_AgeGender$Age_range, levels = c("<18", "18-35", ">35"))
bar_PosAgeGender <- ggplot(Pos_AgeGender, aes(fill = Gender, x = Age_range,
                                               y = Positivity, label = Positivity)) +
                     ylim(0,0.25) +
                     geom_bar(position = "dodge", stat = "identity") +
                     geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
                     geom_text(aes(label = Signif), size = 5, position = position_dodge(0.9),
                               vjust = -1.5) +
                     scale_fill_manual(values = c("lightblue", "navyblue")) +
                     ylab("Pulmonary Embolism Positivity Rate") + 
                     xlab("Age Group") +
                     ggtitle("Pulmonary Embolism Positivity Rate by Age and Gender ")

Pos_BMIGender <- PE_old %>% group_by(BMI_range, Gender) %>% summarise(n = n(),
                                                                      Positivity = mean(Positivity))
Pos_BMIGender$Positivity <- round(Pos_BMIGender$Positivity, 3)
Pos_BMIGender$BMI_range <- factor(Pos_BMIGender$BMI_range, levels = c("<25", "25-40", ">40"))
bar_PosBMIGender <- ggplot(Pos_BMIGender, aes(fill = Gender, x = BMI_range,
                                               y = Positivity, label = Positivity)) + 
                     ylim(0,0.25) +
                     geom_bar(position = "dodge", stat = "identity") +
                     geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
                     scale_fill_manual(values = c("lightblue", "navyblue")) +
                     ylab("Pulmonary Embolism Positivity Rate") + 
                     xlab("BMI Group") +
                     ggtitle("Pulmonary Embolism Positivity Rate by BMI and Gender ")

ggsave(filename = "Images/Bar_PosAgeGender.jpeg", plot = bar_PosAgeGender,
       width = 7, height = 5, device = "jpeg")
ggsave(filename = "Images/Bar_PosBMIGender.jpeg", plot = bar_PosBMIGender,
       width = 7, height = 5, device = "jpeg")

#####-----------------------
## Positivity X location+bmi

Pos_BMILocation <- PE_old %>% group_by(BMI_range, Location) %>% summarise(n = n(),
                                                                      Positivity = mean(Positivity))
Pos_BMILocation$Positivity <- round(Pos_BMILocation$Positivity, 3)
Pos_BMILocation$BMI_range <- factor(Pos_BMI$BMI_range, levels = c("<25", "25-40", ">40"))
bar_PosBMILocation <- ggplot(Pos_BMILocation, aes(fill = Location, x = BMI_range,
                                               y = Positivity, label = Positivity)) + 
                     ylim(0,0.25) +
                     geom_bar(position = "dodge", stat = "identity") +
                     geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
                     scale_fill_manual(values = c("lightblue", "dodgerblue", "mediumblue", "navyblue")) +
                     ylab("Pulmonary Embolism Positivity Rate") + 
                     xlab("BMI Group") +
                     ggtitle("Pulmonary Embolism Positivity Rate by BMI and Location")

ggsave(filename = "Images/Bar_PosBMILocation.jpeg", plot = bar_PosBMILocation,
       width = 7, height = 5, device = "jpeg")

bar_PosLocationBMI <- ggplot(Pos_BMILocation, aes(fill = BMI_range, x = Location,
                                               y = Positivity, label = Positivity)) + 
                     ylim(0,0.25) +
                     geom_bar(position = "dodge", stat = "identity") +
                     geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
                     scale_fill_manual(values = c("lightblue", "dodgerblue", "mediumblue", "navyblue")) +
                     ylab("Pulmonary Embolism Positivity Rate") + 
                     xlab("BMI Group") +
                     labs(fill = "BMI Group") +
                     ggtitle("Pulmonary Embolism Positivity Rate by BMI and Location")

ggsave(filename = "Images/Bar_PosLocationBMI.jpeg", plot = bar_PosLocationBMI,
       width = 7, height = 5, device = "jpeg")

#####--------------------
## Observed Scatter Plots
#####--------------------

## Positivity X raw age

PE_old$Age5 <- cut(PE_old$Age, breaks = seq(0,105, 5), labels = seq(5,105,5), right = FALSE)
PosAge5 <- PE_old %>% group_by(Age5) %>% summarise(Positivity = mean(Positivity))
PosAge5$Age5 <- seq(5,100,5)

scat_PosAge <- ggplot(PosAge5, aes(x = Age5, y = Positivity, color = "dodgerblue")) +
                geom_point(size = 1.5) +
                geom_smooth(method = "lm", se = FALSE,
                                size = 0.5, aes(color = "black")) + 
                geom_label(aes(color = "black"), x = 95, y = 0.08,
                           label = expression(paste(R^2, " = .010"))) +
                scale_colour_manual(values = c("black", "dodgerblue")) + 
                xlab("Age") + ylab("Pulmonary Embolism Positivity Rate") + 
                theme(legend.position = "none") +
                ggtitle("Pulmonary Embolism Positivity Rate by Age")

ggsave(filename = "Images/Scat_PosAge.jpeg", plot = scat_PosAge,
       width = 7, height = 5, device = "jpeg")

## Positivity X raw bmi

PE_old$BMI5 <- cut(PE_old$Age, breaks = seq(10,95, 5), labels = seq(15,95,5), right = FALSE)
PosBMI5 <- PE_old %>% group_by(BMI5) %>% summarise(Positivity = mean(Positivity))
PosBMI5$BMI5 <- seq(15,100,5)

scat_PosBMI <- ggplot(PosBMI5, aes(x = BMI5, y = Positivity, color = "mediumblue")) +
                geom_point(size = 1.5) +
                geom_smooth(method = "lm", se = FALSE,
                                size = 0.5, aes(color = "black")) + 
                geom_label(aes(color = "black"), x = 95, y = 0.08,
                           label = expression(paste(R^2, " = 0.051"))) +
                scale_colour_manual(values = c("black", "mediumblue")) + 
                xlab("BMI") + ylab("Pulmonary Embolism Positivity Rate") + 
                theme(legend.position = "none") +
                ggtitle("Pulmonary Embolism Positivity Rate by BMI")

ggsave(filename = "Images/Scat_PosBMI.jpeg", plot = scat_PosBMI,
       width = 7, height = 5, device = "jpeg")

#####--------------------
## Observed Bar + Pie
#####--------------------

## Number patients by location

Pos_Location <- PE_old %>% group_by(Location) %>% summarise(n = n(), Positivity = mean(Positivity))
Pos_Location$Positivity <- round(Pos_Location$Positivity, 3)

bar_NumLocation <- ggplot(Pos_Location, aes(fill = Location, x = Location,
                                       y = n, label = n)) +
                     geom_bar(position = "dodge", stat = "identity", width = 0.5) +
                     geom_text(size = 3, position = position_dodge(0.5), vjust = -1) +
                     scale_fill_manual(values = c("lightblue", "dodgerblue" ,"mediumblue" ,"navyblue")) +
                     ylab("Number of Patients") + 
                     xlab("Patient Location") +
                     #labs(fill = "Age Group") +
                     ggtitle("Number of Patients by Admittance Location")

ggsave(filename = "Images/Bar_NumLocation.jpeg", plot = bar_NumLocation,
       width = 7, height = 5, device = "jpeg")

Pos_Location$pos <- c(189/2, 189+(534/2), 189+534+(268/2), 189+534+268+(1341/2))
Pos_Location$percent <- paste(round(Pos_Location$n/sum(Pos_Location$n) * 100, digits = 1), "%", sep = "")
Pos_Location$label <- paste(" ", Pos_Location$Location, " \n ", Pos_Location$percent)
Pos_Location$label1 <- c(Pos_Location$label[1], "", Pos_Location$label[3:4])
Pos_Location$label2 <- c("", Pos_Location$label[2], "", "")

pie_NumLocation <- ggplot(Pos_Location, aes(x=factor(1), y=n, fill=Location)) +
                   geom_bar(stat="identity", color = "black") +
                   geom_text(aes(x= factor(1), y=rev(pos), label = label1), size=3.5, vjust = -.1) +
                   geom_text(aes(x= factor(1), y=rev(pos), label = label2), size=3.5) +
                   scale_fill_manual(values = rev(c("lightskyblue2", "lightblue",
                                                    "lightskyblue3","lightskyblue4"))) +
                   theme(axis.ticks = element_blank(), axis.text = element_blank(),
                         axis.title = element_blank(), panel.background = element_blank(),
                         legend.position = "none") +
                   ggtitle("Number of Patients by Admittance Location") +# note y = pos
                   coord_polar(theta = "y")

ggsave(filename = "Images/pie_NumLocation.jpeg", plot = pie_NumLocation,
       width = 7, height = 5, device = "jpeg")
```

```{r}
#####----------------------
## Results Graphics
#####----------------------

## Line Plot of Model Predictions
location <- c("ED", "ED", "ED", "ICU", "ICU", "ICU", "IN", "IN", "IN", "OUT", "OUT", "OUT")
BMI <- rep(c("<25", "25-40", ">40"), 4)
mod_probs <- c(.082, .107, .091, .133, .171, .146, .107, .139, .118, .072, .094, .079)
probs <- data.frame(Location = location, bmi = BMI, probs = mod_probs)
probs$bmi <- factor(probs$bmi, levels = c("<25", "25-40", ">40"))

line_PosPreds <- ggplot(probs, aes(x = bmi, y = probs, group = Location, color = Location)) +
                geom_line(size = .75) + geom_point(size = 1.5) +
                scale_colour_manual(values = c("lightblue", "dodgerblue", "mediumblue", "navyblue")) + 
                xlab("BMI Group") + ylab("Probability of Positive Scan") + 
                ggtitle("Predicted CT Positivity by BMI and Admittance Location")

line_labels_PosPreds <- ggplot(probs, aes(x = bmi, y = probs, group = Location, color = Location,
                                   label = probs)) +
                geom_line(size = .75) + geom_point(size = 1.5) +
                geom_text(size = 3, vjust = -1, color = "black") +
                scale_colour_manual(values = c("lightblue", "dodgerblue", "mediumblue", "navyblue")) + 
                xlab("BMI Group") + ylab("Probability of Positive Scan") + 
                ggtitle("Predicted CT Positivity by BMI and Admittance Location")

ggsave(filename = "Images/Line_PosPreds.jpeg", plot = line_PosPreds,
       width = 7, height = 5, device = "jpeg")
ggsave(filename = "Images/Line_Labels_PosPreds.jpeg", plot = line_labels_PosPreds,
       width = 7, height = 5, device = "jpeg")

Pos_Loc <- PE_old %>% group_by(Location) %>% summarise(n = n(), Positivity = mean(Positivity))
Pos_Loc$Positivity <- round(Pos_Loc$Positivity, 3)
Pos_Loc$Signif1 <- c("*", "**", "*", "**")
Pos_Loc$Signif2 <- c("***", "***", "", "")
bar_PosLoc <- ggplot(Pos_Loc, aes(fill = Location, x = Location, y = Positivity, label = Positivity)) +
                     ylim(0,0.25) +
                     geom_bar(position = "dodge", stat = "identity", width = 0.5) +
                     geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
                     geom_text(aes(label = Signif1), size = 5, position = position_dodge(0.9), 
                               vjust = -1.5) +
                     geom_text(aes(label = Signif2), size = 5, position = position_dodge(0.9),
                               vjust = -2.5) +
                     scale_fill_manual(values = c("lightblue", "dodgerblue", "mediumblue" ,"navyblue")) +
                     ylab("Pulmonary Embolism Positivity Rate") + 
                     xlab("Location") +
                     labs(fill = "Location") +
                     theme(legend.position = "none") +
                     ggtitle("Pulmonary Embolism Positivity Rate by Location")

ggsave(filename = "Images/Bar_PosLocation.jpeg", plot = bar_PosLoc,
       width = 7, height = 5, device = "jpeg")

p1 <- .122; p2 <- .098; n1 <- 1097; n2 <- 1235
pp <- (p1*n1 + p2*n2) / (n1 + n2)
z <- (p1-p2)/(sqrt(pp*(1-pp)*(1/n1 + 1/n2)))
1-pnorm(z)
```

## Fix BMI graph
## Gender positivity (M = .122, F = .098)
## Gender X Age significance, add asterisks
--- 18-35 M v. 18-35 F p = .0295, no other significance
## BMI significance + p-value (chi-square), add asterisks
--- 25-40 higher than <25, p = .024
## Positivity by location, add asterisks, provide caption with p-values
--- ICU > OUT/ED p < .01, IN > ED p = .026
## Fix title of bmi + location
## Give caption for BMI scatterplot